### 2015-05-13 whoop.tex
Feedback on watchdog summarization

To avoid quantifiers and the associated theorem proving burden, we use instead a \emph{watchdog race-checking instrumentation}, adapted from~\cite{bardsley2014engineering}.  Suppose we are analysing entry points $T$ and $U$, and that after translation into Boogie these entry points share a common memory region, $Q$.  When analyzing $T$ and $U$ for races, we introduce an unconstrained symbolic constant $\mathit{watched}_Q$, representing some unspecified index into $Q$; we call this the \emph{watched offset} for $Q$.  We then attempt to prove that it is impossible for $T$ and $U$ to race on $Q$ at index $\mathit{watched}_Q$.  If we can succeed in proving this, we know that $T$ and $U$ must be race-free for the \emph{whole} of $Q$, since the watched offset was arbitrary.  This technique of choosing an arbitrary index to analyse for each map manipulated by an entry point pair can be seen as a form of quantifier elimination: rather than asking the underlying theorem prover to reason for all indices of $Q$, in a quantified manner, we eliminate the quantifier in our encoding, and instead ask the theorem prover to reason about a single, but arbitrary, index of $Q$.

\medskip\noindent\textbf{Watchdog Summarization }
%
Early on during the development of \whoop, we realized that scalability would be a serious issue. When we tried to apply \whoop on the r8169 ethernet driver ($>$7000 lines of C code) with all functions fully inlined, we run out of memory. One reason was that the C to Boogie translation produced hundreds of thousands of Boogie IVL code, which could not fit in memory. Another reason was that some entry points recursively call helper functions. Inherently, recursion does not work well with inlining. To tackle these issues, we developed \emph{watchdog summarization}, a novel summarization technique that allows us to analyze drivers in a scalable fashion and without destroying precision.

Watchdog summarization uses the Houdini~\cite{flanagan2001houdini} invariant inference algorithm (available inside Boogie) to automatically compute summaries (pre- and post-conditions and loop invariants) from a pool of \emph{candidate} invariants. The reason we need loop invariants is that we over-approximate loops using loop-cutting (also available inside Boogie).

>>>>

To avoid quantifiers and the associated theorem proving burden, we use instead a \emph{watchdog race-checking instrumentation}, adapted from~\cite{bardsley2014engineering}.  Suppose we are analysing entry points $T$ and $U$, and that after translation into Boogie these entry points share a common memory region, $\mathit{MR}$.  When analyzing $T$ and $U$ for races, we introduce an unconstrained symbolic constant $\mathit{watched}_{\mathit{MR}}$, representing some unspecified index into $\mathit{MR}$; we call this the \emph{watched offset} for $\mathit{MR}$.  We then attempt to prove that it is impossible for $T$ and $U$ to race on $\mathit{MR}$ at index $\mathit{watched}_{\mathit{MR}}$.  If we can succeed in proving this, we know that $T$ and $U$ must be race-free for the \emph{whole} of $\mathit{MR}$, since the watched offset was arbitrary.  This technique of choosing an arbitrary index to analyse for each map manipulated by an entry point pair can be seen as a form of quantifier elimination: rather than asking the underlying theorem prover to reason for all indices of $\mathit{MR}$, in a quantified manner, we eliminate the quantifier in our encoding, and instead ask the theorem prover to reason about a single, but arbitrary, index of $\mathit{MR}$.

\medskip\noindent\textbf{Watchdog Summarization }
%
Early in the development of \whoop we experimented with reasoning about recursion-free drivers by resorting to full inlining.  We found that this did not scale to large drivers: when we tried to apply \whoop to the r8169 ethernet driver ($>$7000 lines of C code) with all functions fully inlined, we quickly exhausted the memory limits of our experimental platform.  \ADComment{Is it thus the case that r8169 did not exhibit any recursion?}  The C-to-Boogie translation produced, after inlining, hundreds of thousands of lines of Boogie IVL code, consuming a lot of memory and leading to gigantic verification conditions that the Z3 theorem prover could not handle effectively.  \ADComment{I rephrased the previous sentence, because if I remember correctly we did get to the point of invoking Z3.  But maybe I imagined that - perhaps we could not even generate the VC?}  Furthermore, some entry points invoke recursive helper functions, which cannot be fully inlined.

\ADComment{I think the following is the weakest part of this (in general very good) section at the moment.  My concerns are that (a) we have not really invented anything new, we just applied Houdini having guessed candidate preconditions, postconditions and loop invariants based on heuristics, and (b) we have not given much intuition for what these candidates look like.  Could you give an example candidate, using mathematical notation?  For instance, if entry point $T$ has a procedure $f$ that releases mutex $m$, then when instrumenting $T$ with symbol $i$ (using the notation from \S\ref{sec:symbolicpairwise}) we might guess $m \notin \currentlockset{i}$ as a postcondition for $\instrument{f}{i}$.  It would be ideal if you could be as specific as possible about the candidates we generate.}

To make our analysis scalable, and to cater for recursion, we developed \emph{watchdog summarization}, a novel summarization technique that allows us to analyze drivers in a scalable fashion and without destroying precision.
Watchdog summarization uses the Houdini~\cite{flanagan2001houdini} invariant inference algorithm (available inside Boogie) to automatically compute summaries (pre- and post-conditions and loop invariants) from a pool of \emph{candidate} invariants.  The Boogie verifier then automatically checks whether each procedure meets its summary, abstracting each called procedure with its summary in the process; Boogie also checks whether loops respect their invariants and uses the invariants to over-approximate the effects of loops.
